# Project R Functions


## Define Colorschemes

define some slightly less fluorescent versions of currently used colours

```{r echo=TRUE, cache=TRUE}
histological_subtype_palette <- scale_color_manual(values = c(
  "Yolk sac tumour" = "#FEC539",
  "Seminoma/Germinoma" = "#008FFF",
  "Embryonal carcinoma" = "red",
  "Teratoma" = "brown",
  "Gonadal control" = "#00A34E",
  "Testis fibroblasts" = "grey"
))

histological_subtype_palette_wt_ovary_distinction <- scale_color_manual(values = c(
  "Yolk sac tumour" = "#FEC539",
  "Seminoma/Germinoma" = "#008FFF",
  "Embryonal carcinoma" = "red",
  "Teratoma" = "brown",
  "Gonadal control (testis)" = "#00A34E",
  "Gonadal control (ovary)" = "#80d1a7",
  "Testis fibroblasts" = "grey"
))

# make a copy without fibroblasts for plotting cohort-1 (which doesn't have any TF)
histological_subtype_palette_sans_TF <-  scale_color_manual(values = c(
  "Yolk sac tumour" = "#FEC539",
  "Seminoma/Germinoma" = "#008FFF",
  "Embryonal carcinoma" = "red",
  "Teratoma" = "brown",
  "Gonadal control" = "#00A34E"
))

# make a copy of this that will work for plots where its the "fill" not the "colour" attribute used
histological_subtype_palette_fill <-  scale_fill_manual(values = c(
  "Yolk sac tumour" = "#FEC539",
  "Seminoma/Germinoma" = "#008FFF",
  "Embryonal carcinoma" = "red",
  "Teratoma" = "brown",
  "Gonadal control" = "#00A34E",
  "Testis fibroblasts" = "grey"
))

# make a copy without fibroblasts for plotting cohort-1 (which doesn't have any TF)
histological_subtype_palette_fill_sans_TF <-  scale_fill_manual(values = c(
  "Yolk sac tumour" = "#FEC539",
  "Seminoma/Germinoma" = "#008FFF",
  "Embryonal carcinoma" = "red",
  "Teratoma" = "brown",
  "Gonadal control" = "#00A34E"
))
```


## Exclude items from one list in another

This function takes a master vector x and removes items from x which are in excluded_items.

```{r echo=TRUE, cache=TRUE}
exclude <- function(x, excluded_items) {
  x = x[!(x %in% excluded_items)]

  if (any(excluded_items %in% x[!(x %in% excluded_items)])) {
    stop("Error excluding items from list")
  } else {
    return(x)
  }
}
```


## Equalise Annotation/Counts Matrix

This function takes an annotation matrix and a counts matrix and subselects both to only have the cells that are common to both.

```{r echo=TRUE, cache=TRUE}
equalise_annot_counts = function(annotation_matrix, counts_matrix) {
	# make list of intersection of ids between annotation_matrix and counts
	trimmed_ids = colnames(counts_matrix)
	trimmed_ids = rownames(annotation_matrix)[rownames(annotation_matrix) %in% trimmed_ids]
	trimmed_ids = trimmed_ids[trimmed_ids %in% rownames(annotation_matrix)]

	annotation_matrix = annotation_matrix[rownames(annotation_matrix) %in% trimmed_ids,]
	counts_matrix = counts_matrix[,colnames(counts_matrix) %in% trimmed_ids]
	counts_matrix = counts_matrix[,rownames(annotation_matrix)]

	return_tables = list(annotation_matrix, counts_matrix)
	return(return_tables)
}
```


## Trim FeatureCounts Matrix

This function trims a featureCounts matrix to the standard gene x cell matrix.

```{r echo=TRUE, cache=TRUE}
trim_fc_matrix = function(counts_matrix) {
	trimmed_matrix = counts_matrix
	rownames(trimmed_matrix) = trimmed_matrix$Geneid
	trimmed_matrix = trimmed_matrix[,7:dim(trimmed_matrix)[2]]
	return(trimmed_matrix)
}
```



## Deseq2/Volcano Convenience Function

This function takes in arguments of row numbers for group A and group B, as well
as labels for those groups.

Deseq2 is then run to find the differentially expression
between those two groups, and the resulting volcano plot is produced and saved
to the specified subfolder within the "Plots" folder.

Also saved in that folder are TSV and excel spreadsheets. The TSV contains all of
the results from the comparison but the excel document contains just the filtered
set meeting the QC thresholds and then has two tabs to view only the overexpressed
and only the underexpressed subsets.
No filtering is done on the results before plotting or saving.

The function returns a named list with the different outputs.


```{r echo=TRUE, cache=TRUE}
deseq2_plot <- function(type, grpA_label, Aint, grpB_label, Bint, plot_counter, annot_table, raw_counts_matrix, forcenobatch = F, fitType = "parametric", subfolder) {
  if (typeof(grpA_label) != "character") {
    print("ERROR: Unexpected data type (", typeof(grpA_label), ") for argument 'grpA_label'.")
    return(NULL)
  }
  if (typeof(grpB_label) != "character") {
    print("ERROR: Unexpected data type (", typeof(grpB_label), ") for argument 'grpB_label'.")
    return(NULL)
  }
  if (!is.numeric(Aint)) {
    print("ERROR: Unexpected data type (", typeof(Aint), ") for argument 'Aint'.")
    return(NULL)
  }
  if (!is.numeric(Bint)) {
    print("ERROR: Unexpected data type (", typeof(Bint), ") for argument 'Bint'.")
    return(NULL)
  }
  if (!is.data.frame(annot_table)) {
    print("ERROR: Unexpected data type for argument 'annot_table'.")
    return(NULL)
  }
  if (!is.data.frame(raw_counts_matrix)) {
    print("ERROR: Unexpected data type for argument 'raw_counts_matrix'.")
    return(NULL)
  }
  if (!dir.exists(paste0("Plots/", subfolder))) {
    dir.create(paste0("Plots/", subfolder), recursive = TRUE, showWarnings = FALSE)
  }



  plot_counter <- plot_counter + 1
  comparison_title <- paste0(grpB_label, " v ", grpA_label)
  filename <- paste0(plot_counter, ". ", comparison_title, "\n(left side signifies highly expressed in ", grpB_label, ", and right side highly expressed in ", grpA_label, ")")

  deseq_input_annot <- annot_table
  kbl <- kable(deseq_input_annot, "latex", booktabs = T, caption = paste0(plot_counter, ". ", comparison_title)) %>%
    row_spec(Aint, color = "white", background = "#D7261E") %>%
    row_spec(Bint, color = "white", background = "#1ecfd7")
  kbl <- kable_styling(kbl, latex_options = c("scale_down"))
  kable_filename <- paste0("Plots/", subfolder, "/", plot_counter, ". ", comparison_title, " comparison table.pdf")
  save_kable(kbl, file = kable_filename)
  # now do same but not latex
  kbl <- kable(deseq_input_annot, caption = paste0(plot_counter, ". ", comparison_title)) %>%
    row_spec(Aint, color = "white", background = "#D7261E") %>%
    row_spec(Bint, color = "white", background = "#1ecfd7")

  deseq_input_annot[Aint, "CompSide"] <- grpA_label
  deseq_input_annot[Bint, "CompSide"] <- grpB_label
  rownames(deseq_input_annot) <- deseq_input_annot$Row.names
  deseq_input_annot$Row.names <- NULL

  deseq_input_annot <- deseq_input_annot[!is.na(deseq_input_annot$CompSide), ]
  equalised_tables <- equalise_annot_counts(deseq_input_annot, raw_counts_matrix)
  deseq_input_annot <- equalised_tables[[1]]
  trimmed_counts <- equalised_tables[[2]]
  trimmed_counts <- trimmed_counts[, rownames(deseq_input_annot)]
  trimmed_counts <- trimmed_counts + 1

  if ((length(unique(deseq_input_annot$batch)) > 1) & !forcenobatch) {
    deseq_matrix_raw <- suppressMessages(DESeqDataSetFromMatrix(
      countData = trimmed_counts,
      colData = deseq_input_annot,
      design = ~ CompSide + batch
    ))
  } else {
    message(paste0("INFO: Not regressing batch effects as it seems there is only batch:", unique(deseq_input_annot$batch)))
    deseq_matrix_raw <- suppressMessages(DESeqDataSetFromMatrix(
      countData = trimmed_counts,
      colData = deseq_input_annot,
      design = ~CompSide
    ))
  }

  deseq_matrix <- suppressMessages(DESeq(deseq_matrix_raw, fitType = fitType, parallel = F)) # TODO: fix needing parallel=f here in updated R version
  deseq_results <- results(deseq_matrix, contrast = c("CompSide", grpA_label, grpB_label))
  deseq_results <- suppressMessages(lfcShrink(deseq_matrix, parallel = TRUE, type = "ashr", contrast = c("CompSide", grpA_label, grpB_label), res = deseq_results))

  # save tsv with all of the comparison results, not just statistically significant ones
  write.table(x = as.data.frame(deseq_results), paste0("Plots/", subfolder, "/", plot_counter, ". ", "DE results table (unfiltered) ", comparison_title, ".tsv"), sep = "\t", row.names = T)


  # write the excel output of the analysis containing only the statistically significant smallRNAs
  # following request, these now have 3 tabs inside the excel document the first one with
  # everything sorted by adj-p value and then two tabs for overexpressed and underexpressed only
  deseq_filtered_results <- as.data.frame(deseq_results)

  deseq_filtered_results <- deseq_filtered_results[(!is.na(deseq_filtered_results$padj)), ]
  deseq_filtered_results <- deseq_filtered_results[deseq_filtered_results$padj < 0.05, ]
  deseq_filtered_results <- deseq_filtered_results[abs(deseq_filtered_results$log2FoldChange) > 1, ]
  deseq_filtered_results <- deseq_filtered_results[order(deseq_filtered_results$padj), ]

  upreg_filtered_results <- deseq_filtered_results[deseq_filtered_results$log2FoldChange > 0, ]
  upreg_sheet_name <- paste0("Only OE ", nrow(upreg_filtered_results))

  downreg_filtered_results <- deseq_filtered_results[deseq_filtered_results$log2FoldChange < 0, ]
  downreg_sheet_name <- paste0("Only UE ", nrow(downreg_filtered_results))


  excel_filename <- paste0("Plots/", subfolder, "/", plot_counter, ". ", "DE results table ", comparison_title, ".xlsx")
  WriteXLS(
    x = c("deseq_filtered_results", "upreg_filtered_results", "downreg_filtered_results"),
    SheetNames = c(paste0("DE (by adj p-value) ", nrow(deseq_filtered_results)), upreg_sheet_name, downreg_sheet_name),
    ExcelFileName = excel_filename,
    row.names = T
  )


  logFC_plot_filename <- paste0("Plots/", subfolder, "/", plot_counter, ". ", "DE LogFC plot ", comparison_title, ".pdf")
  pdf(file = logFC_plot_filename, width = 7.5, height = 4, title = paste("Log2 FC of all", type, "between", grpA_label, "and", grpB_label))
  plot(deseq_results$log2FoldChange, pch = 16, cex = .9, main = paste("Log2 FC of all", type, "between", grpA_label, "and", grpB_label), ylab = "Log2 Fold Change", col = rgb(red = 0, green = 0, blue = 0, alpha = 0.35))
  dev.off()

  if (interactive()) {
    plot(deseq_results$log2FoldChange, pch = 16, cex = .9, main = paste("Log2 FC of all", type, "between", grpA_label, "and", grpB_label), ylab = "Log2 Fold Change", col = rgb(red = 0, green = 0, blue = 0, alpha = 0.35))
  }

  len_side1 <- length(unique(rownames(deseq_input_annot[deseq_input_annot$CompSide == (unique(deseq_input_annot$CompSide)[1]), ])))
  len_side2 <- length(unique(rownames(deseq_input_annot[deseq_input_annot$CompSide == (unique(deseq_input_annot$CompSide)[2]), ])))
  plot_caption <- paste0(
    "Total features = ", nrow(deseq_results),
    ", Cells in comparison split by '",
    unique(deseq_input_annot$CompSide)[1], "' (n=", len_side1, ") and '", unique(deseq_input_annot$CompSide)[2], "' (n=", len_side2, ")"
  )


  p_cutoff <- 0.05
  lfc_cutoff <- 1
  plt <- EnhancedVolcano::EnhancedVolcano(deseq_results,
    lab = rownames(deseq_results),
    x = "log2FoldChange",
    y = "pvalue",
    pCutoff = p_cutoff,
    FCcutoff = lfc_cutoff,
    labSize = 2.25,
    pointSize = 1.25,
    title = paste0("Differential Expression of ", type, " (p<", p_cutoff, ";lFC>", lfc_cutoff, ")"),
    subtitle = filename,
    caption = plot_caption,
    captionLabSize = 11,
    subtitleLabSize = 11
  )

  volcano_filename <- paste0("Plots/", subfolder, "/", plot_counter, ". Volcano plot ", comparison_title, ".pdf")
  ggsave(plot = plt, filename = volcano_filename, device = "pdf", width = 15, height = 8, dpi = 300)

  # return a named list with items for the counter, serialised volcano ggplot, excel output filename, and serialed html kable comparison table
  results_to_return <- list(counter = plot_counter, volcano = serialize(plt, NULL), excel = excel_filename, comptable = serialize(kbl, NULL))
  return(results_to_return)
}
```


## DE Excel to Heatmap

This function takes a link to an excel spreadsheet output of a differential
expression analysis and plots the statistically significant RNAs as a heatmap.
By default it plots for all samples that are not part of "test-lane", but this is
configurable by specifying a character vector of cell/sample names as the
cells_to_use argument.


```{r echo=TRUE, cache=TRUE}
ExcelHeatmap <- function(excel_path, caption, uponly = FALSE, downonly = FALSE, cells_to_use = testless_cells) {
  t <- read_excel(excel_path)
  t$absLFC <- abs(t$log2FoldChange)
  t <- t[order(t$absLFC, decreasing = T), ]
  t <- t[t$absLFC > 1, ]
  t <- t[!is.na(t$padj), ]
  t <- t[t$padj < 0.05, ]
  colnames(t)[1] <- "piRNA_id"
  t$absLFC <- NULL
  t$pvalue <- NULL
  t$lfcSE <- NULL
  t$baseMean <- NULL

  if (uponly) {
    t <- t[t$log2FoldChange > 0, ]
  }

  if (downonly) {
    t <- t[t$log2FoldChange < 0, ]
  }

  heatmap_mat <- mat[t$piRNA_id, cells_to_use]
  print(Heatmap(heatmap_mat,
    column_labels = annot_table$Hist_Type[annot_table$Row.names %in% colnames(heatmap_mat)],
    column_names_gp = gpar(fontsize = 10),
    row_names_gp = gpar(fontsize = 9),
    top_annotation = HeatmapAnnotation(df = annot_table[annot_table$Row.names %in% colnames(heatmap_mat), ], col = cmplxhmapcols)
  ))
  DT::datatable(t, class = "cell-border stripe", rownames = F, caption = caption)
}
```



## Network Pathway-piRNA Plot

This function takes a _result_df_ dataframe containing a filtered version of the
output returned from the _run_pathfindR()_ function. By filtered, it means removing duplicates
and reducing the number of rows, such as selecting only the top 5 rows of the returned
output. It also takes another dataframe as argument, the _piR_Gene_df_, this is
a dataframe resulting from the merge the original piRNA differential expression
results, with the "flattened" version of those results which has had its piRNAs
renamed to target mRNAs and flattened.


```{r echo=TRUE, cache=TRUE}
piRNA_pathway_network <- function(result_df, piR_Gene_df, comparison_label, num_terms = 10, layout = "stress", use_description = TRUE, node_size = "num_genes") {
  if (!is.numeric(num_terms) & !is.null(num_terms)) {
    stop("`num_terms` must either be numeric or NULL!")
  }
  if (!is.logical(use_description)) {
    stop("`use_description` must either be TRUE or FALSE!")
  }
  ID_column <- ifelse(use_description, "Term_Description",
    "ID"
  )
  val_node_size <- c("num_genes", "p_val")
  if (!node_size %in% val_node_size) {
    stop("`node_size` should be one of ", paste(dQuote(val_node_size),
      collapse = ", "
    ))
  }
  if (!is.data.frame(result_df)) {
    stop("`result_df` should be a data frame")
  }
  necessary_cols <- c(
    ID_column, "lowest_p", "Up_regulated",
    "Down_regulated"
  )
  if (!all(necessary_cols %in% colnames(result_df))) {
    stop(paste(c(
      "All of", paste(necessary_cols, collapse = ", "),
      "must be present in `results_df`!"
    ), collapse = " "))
  }
  if (!is.null(num_terms)) {
    if (nrow(result_df) < num_terms) {
      num_terms <- NULL
    }
  }


  result_df <- result_df[order(result_df$lowest_p, decreasing = FALSE), ]
  if (!is.null(num_terms)) {
    result_df <- result_df[1:num_terms, ]
  }

  graph_df <- data.frame()
  for (i in base::seq_len(nrow(result_df))) {
    up_genes <- unlist(strsplit(
      result_df$Up_regulated[i],
      ", "
    ))
    down_genes <- unlist(strsplit(
      result_df$Down_regulated[i],
      ", "
    ))
    genes <- c(up_genes, down_genes)
    for (gene in genes) {
      graph_df <- rbind(graph_df, data.frame(Term = result_df[
        i,
        ID_column
      ], Gene = gene))
    }
  }
  up_genes <- lapply(result_df$Up_regulated, function(x) {
    unlist(strsplit(
      x,
      ", "
    ))
  })
  up_genes <- unlist(up_genes)

  colnames(piR_Gene_df)[1:2] <- c("Gene", "Term")
  rownames(piR_Gene_df) <- NULL

  graph_df <- rbind(graph_df, piR_Gene_df[, c("Gene", "Term")])


  g <- igraph::graph_from_data_frame(graph_df, directed = FALSE)
  cond_term <- names(igraph::V(g)) %in% result_df[, ID_column]
  cond_up_gene <- names(igraph::V(g)) %in% up_genes

  # label node type
  igraph::V(g)$type <- ifelse(cond_term, "term", "gene")
  cond_pirna <- grepl(pattern = "^piR-", names(igraph::V(g)))
  igraph::V(g)$type[cond_pirna] <- "piRNA"

  # label if gene up or down regulated
  igraph::V(g)$Gene <- ifelse(cond_up_gene, "Up_regulated", "Down_regulated")


  igraph::V(g)$logfc <- NA
  igraph::V(g)$logfc[igraph::V(g)$name %in% piR_Gene_df$Term] <- (-1 * piR_Gene_df$log2FoldChange) # flip LFC again this time so we can see actual piRNA LFC

  igraph::V(g)$p[igraph::V(g)$name %in% piR_Gene_df$Term] <- piR_Gene_df$padj


  set.seed(486)
  ggraph(g, layout = "nicely") +
    geom_edge_link0(edge_colour = "grey66") +
    geom_node_point(data = . %>% filter(type == "term"), color = "black", fill = "#E5D7BF", shape = 24, size = 13) +
    geom_node_point(data = . %>% filter(type == "gene"), aes(colour = Gene), shape = 18, size = 14) + scale_color_manual(values = c("#009E73", "#F8766D"), labels = c("Downregulated in pathway", "Upregulated in pathway")) +
    geom_node_point(data = . %>% filter(type == "piRNA"), aes(fill = logfc, size = abs(logfc)), shape = 21) + scale_fill_distiller(palette = "RdYlBu") +
    geom_node_label(data = . %>% filter(!is.na(type)), ggplot2::aes_(label = ~name), repel = T, size = 4) +
    ggplot2::scale_size(range = c(5, 10))
}
```



## Write Vector Figures to Powerpoint

```{r echo=TRUE, cache=TRUE}
writeFigure <- function(ggplot = NA, plotcmd = NA) {
  if (!missing(ggplot)) {
    plt <- ggplot
  } else if (!missing(plotcmd)) {
    plt <- plotcmd
  }

  if (!file.exists("Figures.ppt")) {
    read_pptx() %>%
      print(target = "Figures.pptx")
  }
  ppt = read_pptx(path = "Figures.pptx")

  add_slide(ppt, layout = "Title and Content", master = "Office Theme") %>%
    ph_with(dml(ggobj = plt), location = ph_location_fullsize()) %>%
    print(target = "Figures.pptx")
}
```
