
# piRNA Pathway Enrichment

## Load predicted mRNA Targets

```{r cache=TRUE, include=FALSE}
i = 0
for (count_file in list.files("Data/piRNA_mRNA_predictions/miRanda_predicted_piRNA_targets", pattern = "*.parsed.txt", full.names = T)) {
	x = read.table(file = count_file, header = F, sep = "\t", stringsAsFactors = F)
	colnames(x)[1] = "piRbaseID"
	colnames(x)[2] = "Transcript"
	colnames(x)[3] = "Score"

	# append each counts file to one another to make a matrix
	if (i == 0) {
		piRNA_predictions = x
	} else {
		piRNA_predictions = rbind(piRNA_predictions, x)
	}

	i = i+1
}
rm(count_file)
rm(x)
rm(i)
```


```{r echo=TRUE}
datatable(head(piRNA_predictions), rownames = F)
```


```{r cache=TRUE, include=FALSE}
if (!file.exists("rds/ensembl_hsapiens.rds")) {
  ensembl_hsapiens <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
  gc() # run garbage collection to free up memory from deleted objects
  saveRDS(ensembl_hsapiens, file = "rds/ensembl_hsapiens.rds")
} else {
  ensembl_hsapiens <- readRDS("rds/ensembl_hsapiens.rds")
}



if (!file.exists("rds/hgnc_symbol_piRNA_prediction.rds")) {
  gc() # run garbage collection to free up memory from deleted objects

  transcript_names <- getBM(
    attributes = c("hgnc_symbol", "refseq_mrna"),
    filters = "refseq_mrna",
    values = piRNA_predictions$Transcript,
    mart = ensembl_hsapiens
  )

  colnames(transcript_names)[2] <- "Transcript"
  piRNA_predictions <- merge(piRNA_predictions, transcript_names, by = "Transcript")
  piRNA_predictions$Gene <- piRNA_predictions$hgnc_symbol
  piRNA_predictions$hgnc_symbol <- NULL
  piRNA_predictions[, c("piRbaseID", "Gene", "Score", "Transcript")]
  saveRDS(piRNA_predictions, file = "rds/hgnc_symbol_piRNA_prediction.rds")
} else {
  piRNA_predictions <- readRDS("rds/hgnc_symbol_piRNA_prediction.rds")
}
```


## Pathway Analysis

### Malignant v Normal

#### mRNA Targets

```{r cache=TRUE, include=FALSE}
deseq_results = list.files("Plots/piRNA DEA/GCT-Subtype Specifics/", pattern = "DE results table \\(unfiltered\\) Healthy v Tumours.tsv$", full.names = T)
deseq_results = read.table(deseq_results)

deseq_results <- deseq_results[!is.na(deseq_results$padj), ]
deseq_results <- deseq_results[deseq_results$padj < 0.05, ]
deseq_results <- deseq_results[abs(deseq_results$log2FoldChange) > 1, ]
deseq_results <- deseq_results[order(abs(deseq_results$log2FoldChange), decreasing = T), ]
deseq_results <- as.data.frame(deseq_results)
deseq_results$piRbaseID <- rownames(deseq_results)
rownames(deseq_results) <- NULL

distinct_piRNA_predictions <- piRNA_predictions[, c("piRbaseID", "Gene")]
distinct_piRNA_predictions <- distinct(distinct_piRNA_predictions)

target_gene_counts <- merge(deseq_results, distinct_piRNA_predictions, by = "piRbaseID")
target_gene_counts <- distinct(target_gene_counts)

if (any(duplicated(target_gene_counts$piRbaseID))) {
  stop("OOh! There are still duplicates in target_gene_counts, identical duplicates were removed so many those that are left are not perfect duplicates, e.g. have different binding sites or scores")
}

gene_map_multi_piRNAs <- target_gene_counts[, c("piRbaseID", "Gene")]
target_gene_counts$piRbaseID <- NULL


target_gene_counts_means <- target_gene_counts %>%
  group_by(Gene) %>%
  summarise_at(c("baseMean", "lfcSE", "pvalue", "padj"), funs(mean))
target_gene_counts_means <- target_gene_counts_means[target_gene_counts_means$Gene != "", ]

target_gene_counts_lfc <- target_gene_counts %>%
  group_by(Gene) %>%
  summarise_at(c("log2FoldChange"), funs(sum))
target_gene_counts_lfc <- target_gene_counts_lfc[target_gene_counts_lfc$Gene != "", ]
rm(target_gene_counts)
target_gene_counts <- merge(target_gene_counts_lfc, target_gene_counts_means, by = "Gene")
rm("target_gene_counts_lfc", "target_gene_counts_means")
target_gene_counts <- as.data.frame(target_gene_counts)

rownames(target_gene_counts) <- target_gene_counts$Gene
target_gene_counts <- target_gene_counts[target_gene_counts$Gene != "", ]
target_gene_counts$Gene <- NULL
target_gene_counts <- target_gene_counts[, c("baseMean", "log2FoldChange", "lfcSE", "pvalue", "padj")]


tumour_mrna <- as.data.frame(target_gene_counts)
rm("deseq_matrix_raw", "deseq_matrix", "deseq_results", "deseq_input_annot", "trimmed_counts", "equalised_tables", "trimmed_counts", "target_gene_counts")

tumour_mrna$TargetGene <- rownames(tumour_mrna)
rownames(tumour_mrna) <- NULL
tumour_mrna$baseMean <- NULL
tumour_mrna$absLFC <- NULL
tumour_mrna$pvalue <- NULL
tumour_mrna$lfcSE <- NULL
tumour_mrna <- tumour_mrna[, c("TargetGene", "log2FoldChange", "padj")]

tumour_mrna <- tumour_mrna[order(abs(tumour_mrna$log2FoldChange), decreasing = T), ]
tumour_mrna$log2FoldChange <- tumour_mrna$log2FoldChange * -1
```



#### KEGG Pathway Targets

```{r}
# Rewrite the manual steps run by the run_pathfindR wrapper to avoid creation of HTML
# result document which conflicts with Rmarkdown compilation
#
#  KEGG_mRNA_targets_malignant_piRNA_DE <- run_pathfindR(
#    input = tumour_mrna,
#    gene_sets = "KEGG",
#    output_dir = "Plots/piRNA-Target Pathway Enrichment/Malignant_v_Normal/",
#    sig_gene_thr = 0,
#    max_to_plot = NULL,
#    pin_name_path = "STRING",
#    visualize_enriched_terms = FALSE,
#    plot_enrichment_chart = FALSE
#  )


if (!file.exists('rds/KEGG_mRNA_targets_malignant_piRNA_DE.rds')) {
  RA_processed <- input_processing(input = tumour_mrna,
                                   pin_name_path  = "STRING")

  # using "BioCarta" as our gene sets for enrichment
  biocarta_list <- fetch_gene_set(gene_sets = "KEGG")
  biocarta_gsets <- biocarta_list[[1]]
  biocarta_descriptions <- biocarta_list[[2]]


  n_iter <- 10 ## number of iterations
  combined_res <- NULL ## to store the result of each iteration

  for (i in 1:n_iter) {

    ###### Active Subnetwork Search
    # output_dir = "Plots/piRNA-Target Pathway Enrichment/Malignant_v_Normal/"
    snws_file <- paste0("active_snws_", i) # Name of output file
    active_snws <- active_snw_search(input_for_search = RA_processed,
                                     pin_name_path = "STRING",
                                     snws_file = snws_file,
                                     sig_gene_thr = 0.00)

    ###### Enrichment Analyses
    current_res <- enrichment_analyses(snws = active_snws,
                                       sig_genes_vec = RA_processed$GENE,
                                       pin_name_path = "STRING",
                                       genes_by_term = biocarta_gsets,
                                       term_descriptions = biocarta_descriptions)

    ###### Combine results via `rbind`
    combined_res <- rbind(combined_res, current_res)
  }

  ###### Summarize Combined Enrichment Results
  summarized_df <- summarize_enrichment_results(combined_res)

  ###### Annotate Affected Genes Involved in Each Enriched Term
  KEGG_mRNA_targets_malignant_piRNA_DE <- annotate_term_genes(result_df = summarized_df,
                                   input_processed = RA_processed,
                                   genes_by_term = biocarta_gsets)

  rm(list = c("summarized_df", "combined_res", "active_snws", "snws_file", "n_iter", "biocarta_descriptions", "biocarta_gsets", "biocarta_list", "RA_processed"))

  gc() # run garbage collection to free up memory from deleted objects
  saveRDS(KEGG_mRNA_targets_malignant_piRNA_DE, file = 'rds/KEGG_mRNA_targets_malignant_piRNA_DE.rds')
  unlink("active_snw_search", recursive = T)
} else {
  KEGG_mRNA_targets_malignant_piRNA_DE = readRDS('rds/KEGG_mRNA_targets_malignant_piRNA_DE.rds')
}
```

```{r}
plt = pathfindR::enrichment_chart(KEGG_mRNA_targets_malignant_piRNA_DE, top_terms = 20)
master_plot_list$figS3$A = plt

ggsave(plt,
       filename = "Plots/piRNA Pathway Enrichment/Top_enriched_Pathways_in_Malignant_GCT.pdf",
       device = "pdf",
       width = 12,
       height = 6)
plt
```


```{r}
datatable(KEGG_mRNA_targets_malignant_piRNA_DE, rownames = F)
```


```{r include=FALSE}
# save table as is for supplementary tables
write.table(KEGG_mRNA_targets_malignant_piRNA_DE,
            file = "Plots/piRNA Pathway Enrichment/Enriched_Pathways_in_Malignant_GCT.tsv",
            sep = "\t",
            row.names = F)
```


##### Top 10

```{r include=FALSE}
# filter to get only those enriched more than |2|
# result_df <- head(KEGG_mRNA_targets_malignant_piRNA_DE[abs(KEGG_mRNA_targets_malignant_piRNA_DE$Fold_Enrichment) > 2, ], 11)
result_df <- head(KEGG_mRNA_targets_malignant_piRNA_DE, 11)
result_df <- result_df[!(result_df$ID %in% c("hsa04211")), ] # this is to remove the duplicate
top_x <- 1:10
result_df <- result_df[top_x, ]


# Merge the target gene DE with piRNA DE results
piR_Gene_df <- data.frame()
for (j in 1:length(result_df$Up_regulated)) {
  set_upreg <- result_df$Up_regulated[[j]]
  set_upreg <- strsplit(set_upreg, split = ", ")[[1]]
  for (i in 1:length(set_upreg)) {
    pirna_in_upreg <- gene_map_multi_piRNAs[gene_map_multi_piRNAs$Gene == set_upreg[i], ]
    pirna_in_upreg <- pirna_in_upreg[!duplicated(pirna_in_upreg$piRbaseID), ]
    piR_Gene_df <- rbind(piR_Gene_df, pirna_in_upreg[, c("Gene", "piRbaseID")])
  }
}
for (j in 1:length(result_df$Down_regulated)) {
  set_downreg <- result_df$Down_regulated[[j]]
  print(set_downreg)
  if (set_downreg != "") {
    set_downreg <- strsplit(set_downreg, split = ", ")[[1]]
    for (i in 1:length(set_downreg)) {
      pirna_in_downreg <- gene_map_multi_piRNAs[gene_map_multi_piRNAs$Gene == set_downreg[i], ]
      pirna_in_downreg <- pirna_in_downreg[!duplicated(pirna_in_downreg$piRbaseID), ]
      piR_Gene_df <- rbind(piR_Gene_df, pirna_in_downreg[, c("Gene", "piRbaseID")])
    }
  }
}

rownames(piR_Gene_df) <- NULL # inherited rownumbers from before merge, the numbers don't mean anything so remove them here
colnames(piR_Gene_df)[1] <- "TargetGene"
piR_Gene_df <- merge(piR_Gene_df, tumour_mrna[, c("TargetGene", "log2FoldChange", "padj")], by = "TargetGene")
piR_Gene_df <- distinct(piR_Gene_df)
rownames(piR_Gene_df) <- NULL
piR_Gene_df <- piR_Gene_df[!duplicated(piR_Gene_df$piRbaseID), ]
datatable(piR_Gene_df, rownames = F)
```


Numerous studies have demonstrated additional roles of piRNA in post-transcriptional gene silencing, targeting mRNAs [@Gou_2014, @Watanabe_2014, @Rouget_2010] in a similar mechanism to miRNA gene repression [@Liu_2019]. This mechanism has previously been been studied in lung carcinoma where introduced piR-55490 reduced cell proliferation through the targeting of mTOR 3′-UTR mRNA [@Peng_2015]. We were thus interested in pathways that piRNAs could be targeting via similar mechanisms.

We attempted to deduce the functional impact of these enriched piRNAs by aligning the list of differentially expressed piRNAs in tumour to the human reference transcriptome. Transcripts matching the quality control thresholds for alignment with our piRNAs were used to construct a pathway enrichment network. The KEGG pathways which saw the greatest enrichment in tumour-derived piRNA targets were mostly related to signalling  pathways (JAK-STAT, RAP1, Calcium, and RAS signalling) and cancer-associated pathways (Prostate cancer, Gastric cancer, non-small cell lung cancer).


```{r fig.width=20, fig.height=12}
plt = piRNA_pathway_network(result_df, piR_Gene_df, comparison_label = "malignant v Normal") +
  ggtitle("Network of Enriched Pathways in GCT predicted from DE piRNAs") +
  theme(plot.title = element_text(size=24))
master_plot_list$fig6 = plt
ggsave(filename = "Plots/piRNA Pathway Enrichment/Pathways1-10.pdf", device = "pdf",width = 10, height = 6, scale = 2)
plt
```

